---
title: "DATASET_COLORSTRIP"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
author: Weiyue Liu^[IOCAS, weiyLiu@outlook.com], Zhongyi Hua^[CACMS, njbxhzy@hotmail.com], Tong Zhou^[IOCAS, tongzhou2017@gmail.com]
vignette: >
  %\VignetteIndexEntry{DATASET_COLORSTRIP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  echo = TRUE,
  message=FALSE,
  warning=FALSE,
  comment = "#>"
)
```

## Introduction
The function of `DATASET_COLORSTRIP` template is to generate corresponding color strips outside the tree for specified tips. The `DATASET_COLORSTRIP` template belongs to the "Text" class (refer to the [Class]() for detail information).

Typically, users add colored strips by entering branch names or nodes, colors (RGB, RGBA, or hexadecimal format), and labels. 

This section shows how to use itol.toolkit to add color strips outside the tree. Without itol.toolkit, users need to manually input colors one by one. This process will be very cumbersome if there are many tips. With the help of itol.toolkit, users only need to import metadata, and the program will automatically generate color information for each tip, greatly improving the efficiency of data processing.

## General process
This section uses [dataset 1](https://github.com/TongZhou2017/itol.toolkit/tree/master/inst/extdata/dataset1) and [dataset 3](https://github.com/TongZhou2017/itol.toolkit/tree/master/inst/extdata/dataset3) as examples to show adding color strips for trees of different sizes (refer to the  [Dataset](https://tongzhou2017.github.io/itol.toolkit/articles/Datasets.html) for detail information). Users can download the data locally from the provided link above. The files are read using a relative path in this document, so please adjust the path based on actual situation.

### Dataset 1
The first step is to load the `newick` format tree file `tree_of_itol_templates.tree` and its corresponding metadata `template_groups`.

The purpose of this part is to add different colored strips for different types of templates.
```{r load data1}
library(itol.toolkit)
library(data.table)
library(ape)
tree_1 <- system.file("extdata",
                      "tree_of_itol_templates.tree",
                       package = "itol.toolkit")
data("template_groups")
```
The input data for `DATASET_COLORSTRIP` contains two columns: The first column is the leaf node name and the second is the group name.We can add color strips to the tips by entering the branch name. 
```{r color strip in tip}
unit_20 <- create_unit(data = template_groups,
                       key = "E020_color_strip_1",
                       type = "DATASET_COLORSTRIP",
                       tree = tree_1)
```
Color strips can also be added through nodes.
```{r color strip in node}
df_data <- data.frame(id = c("tree_sturcture", "theme_style", "text", "basic_plot", "advanced_plot"),
                      data = c("tree_sturcture", "theme_style", "text", "basic_plot", "advanced_plot"))
unit_21 <- create_unit(data = df_data,
                       key = "E021_color_strip_2",
                       type = "DATASET_COLORSTRIP",
                       tree = tree_1)
```
It is worth noting that when adding a color strip through node, the color strip is not displayed outside the tree by default, and can be displayed by turning on `Display internal node values` on the iTOL control panel. At the same time, users can display labels through `Display strip labels` on the iTOL control panel.

### Dataset 3
In actual use, there may be a large number of groups entered by users, which may exceed the number of color plates in some R  packages. In itol.toolkit, automatic generation of up to 75 colors is supported. Tree file `rare-tree.nwk` is a large tree composed of rare OTUs. The `rare-metadata.txt` file stores the abundance information and taxonomy information of rare OTUs.
The purpose of this part is to display taxonomy information at class level of each OTUs through color strips.
```{r load data2}
library(itol.toolkit)
library(data.table)
library(ape)
library(dplyr)
# Users need to `setwd()` to the directory containing dataset3
tree_2 <- read.tree("rare-tree.nwk")
df_metadata <- fread("rare-metadata.txt")
```
Filter the taxonomy information of class in the metadata and generate color strips based on it.
```{r color strip by class}
unit_3 <- create_unit(data = df_metadata %>% 
                        select(ID, Class), 
                      key = "Zheng2022ep_3al_3_strip",
                      type = "DATASET_COLORSTRIP",
                      color = "wesanderson",
                      tree = tree_2)
```
It should be noted that when the tree is large or there are too many annotation information dimensions, it may cause the browser to crash. You should choose another browser or close the browser tab at this time.
