---
title: "DATASET_CONNECTIONS"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
author: Weiyue Liu^[IOCAS, weiyLiu@outlook.com], Zhongyi Hua^[CACMS, njbxhzy@hotmail.com], Tong Zhou^[IOCAS, tongzhou2017@gmail.com]
vignette: >
  %\VignetteIndexEntry{DATASET_CONNECTIONS}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  echo = TRUE,
  message=FALSE,
  warning=FALSE,
  comment = "#>"
)
```

## Introduction
The function of the `DATASET_CONNECTION` template is to draw a line or curve between any two nodes in the tree. Users can customize the appearance of each connection by setting its width, color, style (normal or dashed), and opacity. Additionally, arrowheads can be displayed to indicate the direction of the connection lines. The `DATASET_CONNECTION` template belongs to the "Advanced graphics" class (refer to the [Class]() for detail information).

The nodes are connected by the user through inputting their IDs and specifying the line width, color, style, and label. Without itol.toolkit, the user would have to set the appearance of each connection separately, resulting in a difficult data preparation process for visualization. The itol.toolkit provides consistency to the workflow, enabling the user to perform data processing and visualization preparation seamlessly.The user only needs to input the two nodes being connected, line width, and label text. The program will automatically generate color information, thereby increasing the efficiency of visualization preparation.

## Connection Visualization
This section provides an example of how to draw a connection diagram using [dataset 1](https://github.com/TongZhou2017/itol.toolkit/tree/master/inst/extdata/dataset1) (refer to the [Dataset](https://tongzhou2017.github.io/itol.toolkit/articles/Datasets.html) for detail information) document for detailed information).

The first step is to load the `newick` format tree file `tree_of_itol_templates.tree` and its corresponding metadata `parameter_groups.txt`, `template_groups.txt`. The `parameter_groups.txt` file contains the set of each parameter and the parameters contained in each template. The `template_groups.txt` file contains which class each template belongs to.
```{r load data}
library(itol.toolkit)
library(data.table)
library(dplyr)
library(tidyr)
tree <- system.file("extdata",
                    "tree_of_itol_templates.tree",
                    package = "itol.toolkit")
parameter_groups <- system.file("extdata",
                                "parameter_groups.txt",
                                package = "itol.toolkit")
parameter_groups <- fread(parameter_groups)
data("template_groups")
```
The table used to create the connection should have four columns. The first two columns for the connected points, the third column for the line width, and the fourth column for the labels. In practice, the user needs to prepare the input data according to this format. Here, we simulate the connection between templates by the parameters they contain. The template is stored in the variable `tab_tmp_connect`
```{r data process}
tab_tmp_connect <- parameter_groups[-1, -c(1,2)] %>% 
                   convert_01() %>% 
                   convert_01_to_connect()
tab_tmp_connect$val <- tab_tmp_connect$val -1
tab_tmp_connect <- full_join(tab_tmp_connect,
                             template_groups,
                             by=c("row" = "template"))
tab_tmp_connect <- tab_tmp_connect %>%
                   filter(val > 9) %>%
                   filter(row != col)
unit_32 <- create_unit(data = tab_tmp_connect,
                       key = "E032_connection_1",
                       type = "DATASET_CONNECTION",
                       tree = tree)
```

## Style modification
The maximum line width can be set with the `unit@specific_themes$connection$line$width_max` variable. Arrows can be displayed by setting the `unit@specific_themes$connection$arrow$display` variable to 1. 
