---
title: "DATASET_PIECHART"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
author: Longzhao Li^[IOCAS, 2575760336@qq.com], Zhongyi Hua^[CACMS, njbxhzy@hotmail.com], Tong Zhou^[IOCAS, tongzhou2017@gmail.com]
vignette: >
  %\VignetteIndexEntry{DATASET_PIECHART}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  echo = TRUE,
  message=FALSE,
  warning=FALSE,
  comment = "#>"
)
```

## Introduction
The function of `DATASET_PIECHART` is to draw the pie chart and store drawing information. The `DATASET_PIECHART` template belongs to the "Basic graphics" class (refer to the [Class]() for detail information).

Typically, users cannot save drawing information using iTOL, making reproduce pictures or share them with others difficultly. Here, we provide a convenient way to store **radius** and **the proportion of different groups** of pip charts. This section describes how to itol.toolkit to draw the pie chart and store corresponding information.
 
## General process
This section uses [dataset 1](https://github.com/TongZhou2017/itol.toolkit/tree/master/inst/extdata/dataset1) as an example to show how to draw the line chart. (refer to the  [Dataset](https://tongzhou2017.github.io/itol.toolkit/articles/Datasets.html) for detail information)

### Load data
The first step is to load the `newick` format tree file `tree_of_itol_templates.tree` and its corresponding metadata `df_frequence`. 
```{r load data}
library(itol.toolkit)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)
library(ape)
tree <- system.file("extdata",
                    "tree_of_itol_templates.tree",
                    package = "itol.toolkit")
df_frequence <- system.file("extdata",
                            "templates_frequence.txt",
                             package = "itol.toolkit")
df_frequence <- fread(df_frequence)
names(df_frequence) <- c(
  "id",
  "Li,S. et al. (2022) J. Hazard. Mater.","Zheng,L. et al. (2022) Environ. Pollut.",
  "Welter,D.K. et al. (2021) mSystems",
  "Zhang,L et al. (2022) Nat. Commun.",
  "Rubbens,P. et al. (2019) mSystems",
  "Laidoudi,Y. et al. (2022) Pathogens",
  "Wang,Y. et al. (2022) Nat. Commun.",
  "Ceres,K.M. et al. (2022) Microb. Genomics",
  "Youngblut,N.D. et al. (2019) Nat. Commun.",
  "Balvín,O. et al. (2018) Sci. Rep.",
  "Prostak,S.M. et al. (2021) Curr. Biol.",
  "Dijkhuizen,L.W. et al. (2021) Front. Plant Sci.",
  "Zhang,X. et al. (2022) Microbiol. Spectr.",
  "Peris,D. et al. (2022) PLOS Genet.",
  "Denamur,E. et al. (2022) PLOS Genet.",
  "Dezordi,F.Z. et al. (2022) bioRxiv",
  "Lin,Y. et al. (2021) Microbiome",
  "Wang,Y. et al. (2022) bioRxiv",
  "Qi,Z. et al. (2022) Food Control",
  "Zhou,X. et al. (2022) Food Res. Int.",
  "Zhou,X. et al. (2022) Nat. Commun.")
names(df_frequence) <- stringr::str_remove_all(names(df_frequence),"[()]")
names(df_frequence) <- stringr::str_replace_all(names(df_frequence),",","-")
```
## Data processing and create the unit
Convert wide data to long data. After conversion, the input data fed to `DATASET_PIECHART` 
should have at least two columns: tree tip id and the radius. If there are more than two columns,
If there are more than two columns, the proportion of different groups starts from the third column. The column name is the group name.
```{r}
df_frequence_years <- df_frequence %>% 
  pivot_longer(-id)%>% 
  mutate(years = str_extract(name,"\\d{4}")) %>% 
  replace(is.na(.), 0) %>% 
  group_by(id,years) %>% 
  summarise(value = sum(value)) %>% 
  spread(years,value) %>% 
  replace(is.na(.), 0) %>% 
  mutate(size = log(sum(`2018`,`2019`, `2021`, `2022`)+2)) %>% 
  relocate(size, .after = id)
unit <- create_unit(data = df_frequence_years, 
                    key = "E030_simplebar_2",
                    type = "DATASET_PIECHART",
                    position = 0,
                    tree = tree)
```