---
title: "Installation"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
author: Longzhao Li^[IOCAS, 2575760336@qq.com], Tong Zhou^[IOCAS, tongzhou2017@gmail.com]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  echo = TRUE
)
```


Here is the detailed version of the installation tutorial, the installation process encountered errors, you can find detailed solutions in this article.

If the following solutions do not help you, please leave a comment on the [website](https://github.com/TongZhou2017/itol.toolkit/issues). We will solve the problem as soon as possible and add the solution to this document for more users to read.

## Problems caused by the system

In the Ubuntu system, when installing the `ape` dependency package, there may be the following problems:

```{r message=FALSE, warning=FALSE}
#  error "/usr/bin/ld: cannot find -llapack"
#  error "/usr/bin/ld: cannot find -lgfortran"
```

This is due to a lack of basic tools from the system. Install these tools using shell commands:

```{bash message=FALSE, warning=FALSE}
sudo apt install liblapack-dev libopenblas-dev gfortran
```

## Problems caused by R version

The development environment of `itol.toolkit` package is R4.2.0. Since R version 4.0.0, some important language habit updates have been released, among which the updates of `Rcpp`, rlang and other underlying packages have a great impact on users below R version 4.0.0. `itol.toolkit` running in versions earlier than R4.0.0 may cause some errors caused by underlying dependency packages. Run `itol.toolkit` at R 4.0.0 or later possibly.

### nchar error in quickstart

The following error occurs when running the quickstart tutorial:

```{r message=FALSE, warning=FALSE}
unit <- create_unit(data = df_group,  
                    key = "Quickstart",  
                    type = "DATASET_COLORSTRIP",  
                    tree = tree)

#  Error in char (data[[potential_color]]) :
#  'nchar()' requires a character vector
```



The best way is to install `itol.toolkit` v1.1.2 and later, where we removed the `nchar` dependency in favor of `stringr::str_length`, which is more stable across R versions.

The temporary solution is to explicitly define the color parameter in the [`create_unit`](https://tongzhou2017.github.io/itol.toolkit/reference/create_unit.html) function.
```{r message=FALSE, warning=FALSE}
unit <- create_unit(data = df_group,  
                    key = "Quickstart",  
                    type = "DATASET_COLORSTRIP",  
                    color = "Set1",
                    tree = tree)
```


## Problems caused by dependency packages

R packages from the [Bioconductor](https://www.bioconductor.org/) platform cannot be automatically installed by `install.packages` or `devtools::install_github`. This will result in an error similar to the following:

```{r message=FALSE, warning=FALSE}
#  error "package ‘Biostrings’ is not available (for R version 3.6.3)"
```

The best way is to use [`pak`](https://github.com/r-lib/pak) package to install `itol.toolkit` package. See the [main document](https://tongzhou2017.github.io/itol.toolkit/index.html#installation) for the method.

Another way is to install dependency packages using `BiocManager`:

```{r message=FALSE, warning=FALSE}
# install.packages("BiocManager")
BiocManager::install("Biostrings")
```

The complete dependency package information is as follows. If you have network problems, download the dependency package manually and install it locally. See the [next section](https://tongzhou2017.github.io/itol.toolkit/articles/Installation.html#problems-caused-by-the-network) for details.

| package | platform | function in itol.toolkit |
| --- | --- | --- |
| [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html) | CRAN | Data processing |
| [stringr](https://cran.r-project.org/web/packages/stringr/index.html) | CRAN | String processing; `nchar` repalce |
| [data.table](https://cran.r-project.org/web/packages/data.table/index.html) | CRAN | Data processing |
| [methods](https://www.rdocumentation.org/packages/methods/versions/3.6.2) | CRAN | Object base function |
| [stats](https://www.rdocumentation.org/packages/stats/versions/3.6.2l) | CRAN | Data processing |
| [tidyr](https://www.rdocumentation.org/packages/tidyr/versions/1.3.0) | CRAN | Data frame processing |
| [purrr](https://www.rdocumentation.org/packages/purrr/versions/1.0.1) | CRAN | Data screening |
| [seqinr](https://www.rdocumentation.org/packages/seqinr/versions/4.2-23) | CRAN | Tree file processing |
| [ape](https://cran.r-project.org/web/packages/ape/) | CRAN | Tree file processing |
| [Biostrings](https://www.rdocumentation.org/packages/Biostrings/versions/2.40.2) | Bioconductor | fasta format sequence read |
| [wesanderson](https://www.rdocumentation.org/packages/wesanderson/versions/0.3.6) | CRAN | Color scheme |
| [ggsci](https://cran.r-project.org/web/packages/ggsci/index.html) | CRAN | Color scheme |
| [RColorBrewer](https://cran.r-project.org/web/packages/RColorBrewer/index.html) | CRAN | Color scheme |
| [rstudioapi](https://cran.r-project.org/web/packages/rstudioapi/index.html) | CRAN | RStudio Plug-in base |
| [shiny](https://cran.r-project.org/web/packages/shiny/) | CRAN | RStudio Plug-in base |
| [miniUI](https://cran.r-project.org/web/packages/miniUI/index.html) | CRAN | RStudio Plug-in base |
| [colourpicker](https://cran.r-project.org/web/packages/colourpicker/index.html) | CRAN | RStudio Plug-in base |

## Problems caused by the network

If you encounter similar errors caused by network problems:

```{r message=FALSE, warning=FALSE}
#  Warning in install.packages :
#  unable to access index for repository https://cran.rstudio.com/src/contrib:  
#  cannot open URL 'https://cran.rstudio.com/src/contrib/PACKAGES'  
#  Installing package into ‘C:/Users/25757/Documents/R/win-library/4.0’
#  (as ‘lib’ is unspecified)
#
#  Warning in install.packages :
#  unable to access index for repository https://cran.rstudio.com/src/contrib:  
#  cannot open URL 'https://cran.rstudio.com/src/contrib/PACKAGES'  
#
#  Warning in install.packages :
#  package ‘ape’ is not available (for R version 4.0.0)  
#
#  Warning in install.packages :
#  unable to access index for repository https://cran.rstudio.com/bin/windows/contrib/4.0:  
#  cannot open URL 'https://cran.rstudio.com/bin/windows/contrib/4.0/PACKAGES'
```



Please download the R packages manually from the package homepages in the [previous section](https://tongzhou2017.github.io/itol.toolkit/articles/Installation.html#problems-caused-by-dependency-packages), and then install them locally using the following command:
```{r message=FALSE, warning=FALSE}
install.packages(PathToPkgFile,repos=NULL,type="source")
```
