---
title: "v1.1.10"
author: "Tong Zhou"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Updates:

1. New: Correcting the spelling of the `type` parameter.
2. Update: Beautifying the loading prompt format.
3. Update: Fixing the SHOW_VALUE output parameter for the simplebar unit. 

## 1. Correcting the Spelling of the `type` Parameter

This feature can be used to simplify the process and reduce error rates.
For example, with `DATASET_COLORSTRIP`, it supports partial input, case insensitivity, and recognition of misspellings.

```{r}
library(itol.toolkit)
# read data
tree <- system.file("extdata",
                    "tree_of_itol_templates.tree",
                    package = "itol.toolkit")
data("template_groups")
df_group <- data.frame(id = unique(template_groups$group), 
                       data = unique(template_groups$group))

# 1. Partial input "COLORSTRIP" can be correctly recognized
unit <- create_unit(data = df_group, 
                    key = "Quickstart", 
                    type = "COLORSTRIP", 
                    tree = tree)
# [1] "DATASET_COLORSTRIP"

# 2. Partial input "STRIP" can be correctly recognized
unit <- create_unit(data = df_group, 
                    key = "Quickstart", 
                    type = "STRIP", 
                    tree = tree)
# [1] "DATASET_COLORSTRIP"

# 3. Lowercase input "strip" can be correctly recognized
unit <- create_unit(data = df_group, 
                    key = "Quickstart", 
                    type = "strip", 
                    tree = tree)
# [1] "DATASET_COLORSTRIP"

# 4. Misspelled input "strib" can be correctly recognized
unit <- create_unit(data = df_group, 
                    key = "Quickstart", 
                    type = "strib", 
                    tree = tree)
# [1] "DATASET_COLORSTRIP"
```

## 2. Optimizing Loading Prompts

The previous loading prompts were red, which could mislead users into thinking there was an error. Therefore, the font color has been optimized, but the content still serves as a reminder to properly cite the papers.

```{r}
# You can try running it to see the actual effect
library(itol.toolkit)
# ── Loading ───────────────────────────────────────── itol.toolkit 1.1.9 ──
#
# If you use the itol.toolkit package in published research, please cite:
#
# Zhou, T., Xu, K., Zhao, F., Liu, W., Li, L., Hua, Z., & Zhou, X. (2023).
# itol.toolkit accelerates working with iTOL (Interactive Tree of Life) by an
# automated generation of annotation files. Bioinformatics, 39(6), btad339.
```

## 3. Correcting the SHOW_VALUE Output for Simplebar

An issue was identified during the creation of the `DATASET_SIMPLEBAR` unit, where the output parameter for `value_display` was incorrectly set to `SHOW_VALUES` instead of `SHOW_VALUE`. Setting `unit_3@specific_themes$basic_plot$value_display <- 1` previously generated the field `SHOW_VALUES 1`, but the iTOL web template requires `SHOW_VALUE 1`. The extra "S" in the field name prevented the setting from being properly recognized. After reviewing the `show_value` setting across all templates, it was confirmed that only `DATASET_SIMPLEBAR` requires `SHOW_VALUE`, while other templates continue to use `SHOW_VALUES`.


